---
title: "World Flora"
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# World Flora
This is an attempt to clean up plant names in the trnL and ITS2 OTU files.

## Explore the World Flora R package

Install and load the package
```{r}
#install.packages("WorldFlora")
library(WorldFlora)
library(tidyverse)
```

Download the up-to-date backbone.
```{r}
# download
# WFO.download()

# load into workspace?
WFO.remember(WFO.file = "../classification.csv")
```

Explore the data with an example from the WorldFlora PDF
```{r}
head(WFO.data)

# data from WorldFlora PDF
Sapotaceae <- WFO.data[WFO.data$family == "Sapotaceae",]
Sapotaceae <- Sapotaceae[Sapotaceae$taxonRank == "species", ]
Sapotaceae <- Sapotaceae[Sapotaceae$taxonomicStatus == "Accepted", ]
Sapotaceae <- Sapotaceae[, c("scientificName", "scientificNameAuthorship")]
Sapotaceae <- data.frame(ID = c(1:nrow(Sapotaceae)), Sapotaceae)
names(Sapotaceae)[2:3] <- c("species.name", "author")
head(Sapotaceae)
```

Test out the data with an old plant name: _Panicum arizonicum_ is now _Urochloa arizonica_.
```{r}
# Find Panicum arizonicum records, if they exist
pani_ariz <- WFO.data %>% 
  filter(scientificName == "Panicum arizonicum")
pani_ariz

# get "accepted name" ID
pani_ariz_ID <- pani_ariz %>% select(acceptedNameUsageID)
pani_ariz_ID

updated_name <- semi_join(WFO.data,
                          pani_ariz_ID,
                          by = c("taxonID" = "acceptedNameUsageID")) %>% 
  select(scientificName)
updated_name
```

It works! How would we expand this into a functional for loop?

Things to consider:
  
  - whether name in OTU/WTU csv has synonyms (might not need this if written correctly...)
  - whether name in OTU/WTU csv is "Accepted" or not
  - species names in OTU/WTU need to be cleaned (separators, varietals, etc.)
  - what to do with plants with no species-level id
```{r}

```
  
  